---
title: "Multiple Correspondance Analysis"
author: "Richa Singh"
date: "10/10/2018"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
rm(list = ls())
graphics.off()
```

## Multiple Correspondance Analysis 

Multiple correspondence analysis (MCA) is an extension of correspondence analysis (CA) which allows one to analyze the pattern of relationships of several categorical dependent variables. MCA is used to analyze a set of observations described by a set of nominal variables. Each nominal variable comprises several levels, and each of these levels is coded as a binary variable. MCA can also accommodate quantitative variables by recoding them as bins.Because MCA has been (re)discovered many times, equivalent methods are known under several different names such as optimal scaling, optimal or appropriate scoring, dual scaling, homogeneity analysis, scalogram analysis, and quantification method.
Technically MCA is obtained by using a standard correspondence analysis on an indicator matrix (a matrix whose entries are 0 or 1). The percentages of explained variance need to be corrected, and the correspondence analysis interpretation of inter-point distances needs to be adapted.

## Dataset 

The dataset consists of 1233 observations and 32 variables describing the HR-IBM Employees. The variables Sub,Age,Monthly Income,Daily Rate,Hourly Rate, MonthlyRate,DistanceFromeHome,PerformanceRating, Education, JobInvolvement, Joblevel,StockOptionLevel,PercentSalaryHike,TotalWorkingYears,TrainingTimesLastYear,WorkLifeBalance,WorkLifeBalance,YearsAtCompany,YearsInCurrentRole,YearsSinceLastPromotion,YearsWithCurrManager,EnvironmentSatisfaction,JobSatisfaction,RelationshipSatisfaction,Attrition,BusinessTravel,EducationField,Gender,JobRole,MaritalStatus,OverTime are Qualitative variable.

```{r}
my_data <- read.csv("IBM-HR-Emplyee-NoAttrition.csv")
cols <- colnames(my_data)
rownames(my_data) <- my_data$Subj
data1 <- my_data[,11:32]
data1$Age <- my_data$Age
```

## Distribution of the quantitative data 

```{r}
col_no = 1
par(mfrow=c(2,2)) 
for (val in colnames(data1)){
  hist(data1[,col_no],breaks = 10, main=val, xlab = "")
  col_no = col_no +1
}
```

## Binning the quantitative data

```{r}
hist(data1$MonthlyIncome) 
abline(v = c(5000,10000), col = "red", lwd =2)
```

```{r}
# Binning of Monthly Income 
data1$MonthlyIncome <- cut(data1$MonthlyIncome,breaks = c(min(data1$MonthlyIncome)-1,5000 ,10000, max(data1$MonthlyIncome)+1),labels = c(1,2,3))
cor(as.numeric(data1$MonthlyIncome),my_data$MonthlyIncome,method = "spearman")
```

```{r}
hist(data1$DailyRate) 
abline(v = c(600,1100), col = "red", lwd =2)
```


```{r}
data1$DailyRate <- cut(data1$DailyRate,breaks = c(min(data1$DailyRate)-1,600 ,1100, max(data1$DailyRate)+1),labels = c(1,2,3))
cor(as.numeric(data1$DailyRate),my_data$DailyRate,method = "spearman")
```
```{r}
hist(data1$HourlyRate) 
abline(v = c(55,80), col = "red", lwd =2)
```

```{r}
#Binning hourly rate
data1$HourlyRate <- cut(data1$HourlyRate,breaks = c(min(data1$HourlyRate)-1,55 ,80, max(data1$HourlyRate)+1),labels = c(1,2,3))
cor(as.numeric(data1$HourlyRate),my_data$HourlyRate,method = "spearman")
```

```{r}
hist(data1$MonthlyRate) 
abline(v = c(10000,17500), col = "red", lwd =2)
```

```{r}
#Binning Monthly rate 
data1$MonthlyRate <- cut(data1$MonthlyRate,breaks = c(min(data1$MonthlyRate)-1,10000 ,17500, max(data1$MonthlyRate)+1),labels = c(1,2,3))
cor(as.numeric(data1$MonthlyRate),my_data$MonthlyRate,method = "spearman")
```
```{r}
hist(data1$DistanceFromHome) 
abline(v = c(6), col = "red", lwd =2)
```

```{r}
#Binning Distance From Home  
data1$DistanceFromHome <- cut(data1$DistanceFromHome,breaks = c(min(data1$DistanceFromHome)-1,6 , max(data1$DistanceFromHome)+1),labels = c(1,2))
cor(as.numeric(data1$DistanceFromHome),my_data$DistanceFromHome,method = "spearman")
```

```{r}
hist(data1$PerformanceRating) 
abline(v = c(3.5), col = "red", lwd =2)
```

```{r}
hist(data1$Education) 
abline(v = c(2,3), col = "red", lwd =2)
```


```{r}
#Binning for Education
data1$Education <- cut(data1$Education,breaks = c(min(data1$Education)-1, 2,3 , max(data1$Education)+1),labels = c(1,2,3))
cor(as.numeric(data1$Education),my_data$Education,method = "spearman")
```

```{r}
hist(data1$JobInvolvement) 
abline(v = c(2.5,3.5), col = "red", lwd =2)
```

```{r}
#Binning for Job Involvement 
data1$JobInvolvement <- cut(data1$JobInvolvement,breaks = c(min(data1$JobInvolvement)-1,2.5,3.5, max(data1$JobInvolvement)+1),labels = c(1,2,3))
cor(as.numeric(data1$JobInvolvement),my_data$JobInvolvement,method = "spearman")
```

```{r}
hist(data1$JobLevel) 
abline(v = c(2,4), col = "red", lwd =2)
```

```{r}
#Binning JobLevel
data1$JobLevel <- cut(data1$JobLevel,breaks = c(min(data1$JobLevel)-1,2,4, max(data1$JobLevel)+1),labels = c(1,2,3))
cor(as.numeric(data1$JobLevel),my_data$JobLevel,method = "spearman")
```

```{r}
hist(data1$StockOptionLevel) 
abline(v = c(0.5,1.5,2.5), col = "red", lwd =2)
```
Leaving the Stock Option Level as it is 

```{r}
hist(data1$NumCompaniesWorked) 
abline(v = c(2,6), col = "red", lwd =2)
```

```{r}
#Binning Number of Companies Worked 
data1$NumCompaniesWorked <- cut(data1$NumCompaniesWorked,breaks = c(min(data1$NumCompaniesWorked)-1, 2,6, max(data1$NumCompaniesWorked)+1),labels = c(1,2,3))
cor(as.numeric(data1$NumCompaniesWorked),my_data$NumCompaniesWorked,method = "spearman")
```

```{r}
hist(data1$TotalWorkingYears) 
abline(v = c(10), col = "red", lwd =2)
```

```{r}
## Binning total working years
data1$TotalWorkingYears <- cut(data1$TotalWorkingYears,breaks = c(min(data1$TotalWorkingYears)-1, 10, max(data1$TotalWorkingYears)+1),labels = c(1,2))
cor(as.numeric(data1$TotalWorkingYears),my_data$TotalWorkingYears,method = "spearman")
```

```{r}
hist(data1$PercentSalaryHike) 
abline(v = c(13,19), col = "red", lwd =2)
```

```{r}
## Binning percentage Salary Hike
data1$PercentSalaryHike <- cut(data1$PercentSalaryHike,breaks = c(min(data1$PercentSalaryHike)-1, 13,19, max(data1$PercentSalaryHike)+1),labels = c(1,2,3))
cor(as.numeric(data1$PercentSalaryHike),my_data$PercentSalaryHike,method = "spearman")

```
```{r}
hist(data1$WorkLifeBalance) 
abline(v = c(2), col = "red", lwd =2)
```

```{r}
## Binning WorkLifeBalance
data1$WorkLifeBalance <- cut(data1$WorkLifeBalance,breaks = c(min(data1$WorkLifeBalance)-1, 2, max(data1$WorkLifeBalance)+1),labels = c(1,2))
cor(as.numeric(data1$PercentSalaryHike),my_data$PercentSalaryHike,method = "spearman")

```

```{r}
hist(data1$Age) 
abline(v = c(30,40), col = "red", lwd =2)
```

```{r}
## Binning Age 
data1$Age <- cut(data1$Age,breaks = c(min(data1$Age)-1, 30,40, max(data1$Age)+1),labels = c(1,2,3))
cor(as.numeric(data1$Age),my_data$Age,method = "spearman")
```
```{r}
hist(data1$TrainingTimesLastYear) 
abline(v = c(2,4), col = "red", lwd =2)
```

```{r}
## Binning TrainingTimesLastYear
data1$TrainingTimesLastYear <- cut(data1$TrainingTimesLastYear,breaks = c(min(data1$TrainingTimesLastYear)-1, 2,4, max(data1$TrainingTimesLastYear)+1),labels = c(1,2,3))
cor(as.numeric(data1$TrainingTimesLastYear),my_data$TrainingTimesLastYear,method = "spearman")

```

```{r}
hist(data1$YearsAtCompany) 
abline(v = c(5,15), col = "red", lwd =2)
```

```{r}
## Binning Years at Company 
data1$YearsAtCompany <- cut(data1$YearsAtCompany,breaks = c(min(data1$YearsAtCompany)-1, 5,15, max(data1$YearsAtCompany)+1),labels = c(1,2,3))
cor(as.numeric(data1$YearsAtCompany),my_data$YearsAtCompany,method = "spearman")

```

```{r}
hist(data1$YearsInCurrentRole) 
abline(v = c(4,10), col = "red", lwd =2)
```

```{r}
## Binning Years in Current Role 
data1$YearsInCurrentRole <- cut(data1$YearsInCurrentRole,breaks = c(min(data1$YearsInCurrentRole)-1, 4,10, max(data1$YearsInCurrentRole)+1),labels = c(1,2,3))
cor(as.numeric(data1$YearsInCurrentRole),my_data$YearsInCurrentRole,method = "spearman")

```

```{r}
hist(data1$YearsSinceLastPromotion) 
abline(v = c(2,7), col = "red", lwd =2)
```

```{r}
## Binning Years since last Promotion
data1$YearsSinceLastPromotion <- cut(data1$YearsSinceLastPromotion,breaks = c(min(data1$YearsSinceLastPromotion)-1, 2,7, max(data1$YearsSinceLastPromotion)+1),labels = c(1,2,3))
cor(as.numeric(data1$YearsSinceLastPromotion),my_data$YearsSinceLastPromotion,method = "spearman")

```

```{r}
hist(data1$YearsWithCurrManager) 
abline(v = c(4,10), col = "red", lwd =2)
```

```{r}
## Binning Years with current Manager 
data1$YearsWithCurrManager <- cut(data1$YearsWithCurrManager,breaks = c(min(data1$YearsWithCurrManager)-1, 4,10, max(data1$YearsWithCurrManager)+1),labels = c(1,2,3))
cor(as.numeric(data1$YearsWithCurrManager),my_data$YearsWithCurrManager,method = "spearman")

```

```{r}
hist(data1$JobSatisfaction) 
abline(v = c(1.5,2,3,4), col = "red", lwd =2)
```
```{r}
hist(data1$RelationshipSatisfaction) 
abline(v = c(1.5,2,3,4), col = "red", lwd =2)
```
```{r}
hist(data1$EnvironmentSatisfaction) 
abline(v = c(1.5,2,3,4), col = "red", lwd =2)
```
## Heat Map 

```{r}
library(corrplot)
data2 <- my_data[,11:32]
data2$Age <- my_data$Age
cor.my_data <- cor(data2)
corrplot(cor.my_data, method = "ellipse")
```


```{r}
library(data4PCCAR)
library(PTCA4CATA)
library(corrplot)
library(ggplot2)
library(ExPosition)
library(InPosition)
# Heat Map for Burt Table
#heatMapIJ <- makeggHeatMap4CT(data2,colorAttributes = col4J, fontSize.x = 15)
#print(heatMapIJ)
```

```{r}
str(data1)
```


```{r}
data1$PerformanceRating <- as.factor(data1$PerformanceRating)
data1$JobInvolvement <- as.factor(data1$JobInvolvement)
data1$StockOptionLevel <- as.factor(data1$StockOptionLevel)
data1$EnvironmentSatisfaction <- as.factor(data1$EnvironmentSatisfaction)
data1$JobSatisfaction <- as.factor(data1$JobSatisfaction)
data1$RelationshipSatisfaction <- as.factor(data1$RelationshipSatisfaction)
```

## Scree plot+ Inference Results

```{r}

resMCA.sym <- epMCA(data1 ,make_data_nominal = TRUE ,DESIGN = my_data$Department ,make_design_nominal = TRUE,graphs = FALSE, symmetric = TRUE)

resMCA.asym <- epMCA(data1 ,make_data_nominal = TRUE ,DESIGN = my_data$Department ,make_design_nominal = TRUE,graphs = FALSE, symmetric = FALSE)

resMCA.inf <- epMCA.inference.battery(data1, make_data_nominal = TRUE, DESIGN = my_data$Department ,make_design_nominal =TRUE,graphs = FALSE)

```

```{r}

resMCA.sym1 <- epMCA(data1 ,make_data_nominal = TRUE ,DESIGN = my_data$Gender ,make_design_nominal = TRUE,graphs = FALSE, symmetric = TRUE)

resMCA.asym1 <- epMCA(data1 ,make_data_nominal = TRUE ,DESIGN = my_data$Gender ,make_design_nominal = TRUE,graphs = FALSE, symmetric = FALSE)

resMCA.inf1 <- epMCA.inference.battery(data1, make_data_nominal = TRUE, DESIGN = my_data$Gender ,make_design_nominal =TRUE,graphs = FALSE)

```


```{r}
PlotScree(ev = resMCA.sym$ExPosition.Data$eigs, 
          p.ev =  resMCA.inf$Inference.Data$components$p.vals,
          title = 'IBM-No-Attririon data Set. Eigenvalues Inference',
          plotKaiser = TRUE
)
```

## Factor Maps for I for symmetrical plot for component 1 and 2 

Symmetric plot deals with the average comparison 
```{r}

constraints.mca <- minmaxHelper(mat1 = resMCA.sym$ExPosition.Data$fi, mat2  = resMCA.sym$ExPosition.Data$fj)
constraints.asym <- minmaxHelper(mat1 = resMCA.asym$ExPosition.Data$fi , mat2  = resMCA.asym$ExPosition.Data$fi)

color <- prettyGraphsColorSelection(NCOL(data1))
baseMap.i1 <- createFactorMap(resMCA.sym$ExPosition.Data$fi,constraints = constraints.mca,
                            col.points = resMCA.sym$Plotting.Data$fi.col, axis1 = 1, axis2 = 2,
                            cex = 1, pch = 20,
                            display.labels = FALSE
)
label4Map1 <- createxyLabels.gen(1,2,
                                lambda =resMCA.sym$ExPosition.Data$eigs,
                                tau = resMCA.sym$ExPosition.Data$t)

a1 <- baseMap.i1$zeMap + baseMap.i1$zeMap_dots +label4Map1
print(a1)

```

## Factor Maps for I for symmetrical plot for component 2 and 3 


```{r}

color <- prettyGraphsColorSelection(NCOL(data1))
baseMap.i2 <- createFactorMap(resMCA.sym$ExPosition.Data$fi,constraints = constraints.mca,
                            col.points = resMCA.sym$Plotting.Data$fi.col, axis1 = 2, axis2 = 3,
                            cex = 1, pch = 20,
                            display.labels = FALSE
)
label4Map2 <- createxyLabels.gen(1,2,
                                lambda =resMCA.sym$ExPosition.Data$eigs,
                                tau = resMCA.sym$ExPosition.Data$t)

a2 <- baseMap.i2$zeMap + baseMap.i2$zeMap_dots +label4Map2
print(a2)

```

## Factor Maps for I for symmetrical plot for component 3 and 4 


```{r}

color <- prettyGraphsColorSelection(NCOL(data1))
baseMap.i3 <- createFactorMap(resMCA.sym$ExPosition.Data$fi,constraints = constraints.mca,
                            col.points = resMCA.sym$Plotting.Data$fi.col, axis1 = 3, axis2 = 4,
                            cex = 1, pch = 20,
                            display.labels = FALSE
)
label4Map3 <- createxyLabels.gen(1,2,
                                lambda =resMCA.sym$ExPosition.Data$eigs,
                                tau = resMCA.sym$ExPosition.Data$t)

a3 <- baseMap.i3$zeMap + baseMap.i3$zeMap_dots +label4Map3
print(a3)

```

## Factor Maps for J for symmetrical plot for component 1

```{r}
baseMap.j1 <- createFactorMap(resMCA.sym$ExPosition.Data$fj, constraints = constraints.mca,
                             color.points = "black",text.cex = 2,cex = 1,pch = 21, axis1 = 1, axis2 = 2)

print(baseMap.j1$zeMap)
dev.new()
print(baseMap.i1$zeMap + baseMap.j1$zeMap_dots + baseMap.j1$zeMap_text)

```

## Factor Maps for J for symmetrical plot for component 2

```{r}
baseMap.j2 <- createFactorMap(resMCA.sym$ExPosition.Data$fj, constraints = constraints.mca,
                             color.points = "black",text.cex = 2,cex = 1,pch = 21, axis1 = 2, axis2 = 3)

print(baseMap.j2$zeMap)
dev.new()
print(baseMap.i2$zeMap + baseMap.j2$zeMap_dots + baseMap.j2$zeMap_text)

```

## Factor Maps for J for symmetrical plot for component 3

```{r}
baseMap.j3 <- createFactorMap(resMCA.sym$ExPosition.Data$fj, constraints = constraints.mca,
                             color.points = "black",text.cex = 2,cex = 1,pch = 21, axis1 = 3,axis2 = 4)

print(baseMap.j3$zeMap)
dev.new()
print(baseMap.i3$zeMap + baseMap.j3$zeMap_dots + baseMap.j3$zeMap_text)

```

## Biplot for symmetrical plot for component 1 and 2 for Department datatype

```{r}
col4J1 <- prettyGraphsColorSelection(NCOL(data1))
symMap1  <- createFactorMapIJ(resMCA.sym$ExPosition.Data$fi,resMCA.sym$ExPosition.Data$fj,
                             col.points.i = resMCA.sym$Plotting.Data$fi.col,
                             col.points.j = "black",
                             col.labels.i = resMCA.sym$Plotting.Data$fi.col ,
                             col.labels.j = "black" ,
                            cex.i = 2.5, pch.i = 20, 
                             pch.j = 21,  cex.j = 2.5,text.cex.j =2, axis1 = 1,axis2 = 2, title = "Symmetrical plot for components 1 and 2 for Department Type",
                             alpha.axes = 0.2,alpha.points.i = 1)


labels4MCA1 <- createxyLabels(resCA = resMCA.sym, x_axis = 1,y_axis = 2)

map.IJ.sym1 <- symMap1$baseMap + symMap1$I_points 
map.IJ.sym2 <- symMap1$baseMap + symMap1$J_labels + symMap1$J_points + labels4MCA1
map.IJ.sym3 <- symMap1$baseMap + symMap1$I_points +
  symMap1$J_labels + symMap1$J_points + labels4MCA1
print(map.IJ.sym1)
print(map.IJ.sym2)
print(map.IJ.sym3)

```

## Biplot for symmetrical plot for component 1 and 2 for Gender datatype

```{r}
col4J1 <- prettyGraphsColorSelection(NCOL(data1))
symMap1  <- createFactorMapIJ(resMCA.sym1$ExPosition.Data$fi,resMCA.sym1$ExPosition.Data$fj,
                             col.points.i = resMCA.sym1$Plotting.Data$fi.col,
                             col.points.j = "black",
                             col.labels.i = resMCA.sym1$Plotting.Data$fi.col ,
                             col.labels.j = "black" ,
                            cex.i = 2.5, pch.i = 20, 
                             pch.j = 21,  cex.j = 2.5,text.cex.j =2, axis1 = 1,axis2 = 2, title = "Symmetrical plot for components 1 and 2 for Gender Type",
                             alpha.axes = 0.2,alpha.points.i = 1)


labels4MCA1 <- createxyLabels(resCA = resMCA.sym1, x_axis = 1,y_axis = 2)

map.IJ.sym1 <- symMap1$baseMap + symMap1$I_points 
map.IJ.sym2 <- symMap1$baseMap + symMap1$J_labels + symMap1$J_points + labels4MCA1
map.IJ.sym3 <- symMap1$baseMap + symMap1$I_points +
  symMap1$J_labels + symMap1$J_points + labels4MCA1
print(map.IJ.sym1)
print(map.IJ.sym2)
print(map.IJ.sym3)

```

## Biplot for symmetrical plot for component 2 and 3 for Department datatype

```{r}
symMap2  <- createFactorMapIJ(resMCA.sym$ExPosition.Data$fi,resMCA.sym$ExPosition.Data$fj,
                             col.points.i = resMCA.sym$Plotting.Data$fi.col,
                             col.points.j = "black",
                             col.labels.i = resMCA.sym$Plotting.Data$fi.col ,
                             col.labels.j = "black" ,
                            cex.i = 5, pch.i = 20, 
                             pch.j = 21,  text.cex.j =2, axis1 = 2,axis2 = 3, title = "Symmetrical plot for components 2 and 3 for Department Type",
                             alpha.axes = 0.2,alpha.points.i = 1)

labels4MCA2 <- createxyLabels(resCA = resMCA.sym, x_axis = 2, y_axis = 3)

map.IJ.sym11 <- symMap2$baseMap + symMap2$I_points 
map.IJ.sym21 <- symMap2$baseMap + symMap2$J_labels + symMap2$J_points + labels4MCA2
map.IJ.sym31 <- symMap2$baseMap + symMap2$I_points +
  symMap2$J_labels + symMap2$J_points + labels4MCA2
print(map.IJ.sym11)
print(map.IJ.sym21)
print(map.IJ.sym31)

```

## Biplot for symmetrical plot for component 2 and 3 for Gender datatype

```{r}
symMap2  <- createFactorMapIJ(resMCA.sym1$ExPosition.Data$fi,resMCA.sym1$ExPosition.Data$fj,
                             col.points.i = resMCA.sym1$Plotting.Data$fi.col,
                             col.points.j = "black",
                             col.labels.i = resMCA.sym1$Plotting.Data$fi.col ,
                             col.labels.j = "black" ,
                            cex.i = 5, pch.i = 20, 
                             pch.j = 21,  text.cex.j =2, axis1 = 2,axis2 = 3, title = "Symmetrical plot for components 2 and 3 for Gender Type",
                             alpha.axes = 0.2,alpha.points.i = 1)

labels4MCA2 <- createxyLabels(resCA = resMCA.sym1, x_axis = 2, y_axis = 3)

map.IJ.sym11 <- symMap2$baseMap + symMap2$I_points 
map.IJ.sym21 <- symMap2$baseMap + symMap2$J_labels + symMap2$J_points + labels4MCA2
map.IJ.sym31 <- symMap2$baseMap + symMap2$I_points +
  symMap2$J_labels + symMap2$J_points + labels4MCA2
print(map.IJ.sym11)
print(map.IJ.sym21)
print(map.IJ.sym31)

```

## Biplot for symmetrical plot for component 1 and 3 for Department datatype

```{r}
symMap3  <- createFactorMapIJ(resMCA.sym$ExPosition.Data$fi,resMCA.sym$ExPosition.Data$fj,
                             col.points.i = resMCA.sym$Plotting.Data$fi.col,
                             col.points.j = "black",
                             col.labels.i = resMCA.sym$Plotting.Data$fi.col ,
                             col.labels.j = "black" ,
                            cex.i = 5, pch.i = 20, 
                             pch.j = 21,  text.cex.j =2, axis1 = 1,axis2 = 3, title = "Symmetrical plot for components 1 and 3 for Department Type",
                             alpha.axes = 0.2,alpha.points.i = 1)

labels4MCA3 <- createxyLabels(resCA = resMCA.sym, x_axis = 1,y_axis = 3)

map.IJ.sym31 <- symMap3$baseMap + symMap3$I_points 
map.IJ.sym32 <- symMap3$baseMap + symMap3$J_labels + symMap3$J_points + labels4MCA3
map.IJ.sym33 <- symMap3$baseMap + symMap3$I_points +
  symMap3$J_labels + symMap3$J_points + labels4MCA3
print(map.IJ.sym31)
print(map.IJ.sym32)
print(map.IJ.sym33)

```

## Biplot for symmetrical plot for component 1 and 3 for Gender datatype

```{r}
symMap3  <- createFactorMapIJ(resMCA.sym1$ExPosition.Data$fi,resMCA.sym1$ExPosition.Data$fj,
                             col.points.i = resMCA.sym1$Plotting.Data$fi.col,
                             col.points.j = "black",
                             col.labels.i = resMCA.sym1$Plotting.Data$fi.col ,
                             col.labels.j = "black" ,
                            cex.i = 5, pch.i = 20, 
                             pch.j = 21,  text.cex.j =2, axis1 = 1,axis2 = 3, title = "Symmetrical plot for components 1 and 3 for Gender Type",
                             alpha.axes = 0.2,alpha.points.i = 1)

labels4MCA3 <- createxyLabels(resCA = resMCA.sym1, x_axis = 1,y_axis = 3)

map.IJ.sym31 <- symMap3$baseMap + symMap3$I_points 
map.IJ.sym32 <- symMap3$baseMap + symMap3$J_labels + symMap3$J_points + labels4MCA3
map.IJ.sym33 <- symMap3$baseMap + symMap3$I_points +
  symMap3$J_labels + symMap3$J_points + labels4MCA3
print(map.IJ.sym31)
print(map.IJ.sym32)
print(map.IJ.sym33)

```

## Factor Maps for I for asymmetrical plot 

```{r}

color <- prettyGraphsColorSelection(NCOL(data1))
baseMap.i <- createFactorMap(resMCA.asym$ExPosition.Data$fi,constraints = constraints.asym,
                            col.points = resMCA.asym$Plotting.Data$fi.col,
                            cex = 1, pch = 20,
                            display.labels = FALSE
)
label4Map <- createxyLabels.gen(1,2,
                                lambda =resMCA.asym$ExPosition.Data$eigs,
                                tau = resMCA.asym$ExPosition.Data$t)

a <- baseMap.i$zeMap + baseMap.i$zeMap_dots +label4Map
print(a)

```

## Factor Maps for J for asymmetrical plot 

```{r}
baseMap.j <- createFactorMap(resMCA.asym$ExPosition.Data$fj, constraints = constraints.asym,
                             color.points = "black",text.cex = 2,cex = 1,pch = 21)

print(baseMap.j$zeMap)
dev.new()
print(baseMap.i$zeMap + baseMap.j$zeMap_dots + baseMap.j$zeMap_text)

```

## Biplot for asymmetrical plot for component 1

Asymmetric plot deals with the attraction by the variables 
```{r}
asymMap1  <- createFactorMapIJ(resMCA.asym$ExPosition.Data$fi,resMCA.asym$ExPosition.Data$fj,
                             col.points.i = resMCA.asym$Plotting.Data$fi.col,
                             col.labels.j = resMCA.asym$Plotting.Data$fj.col,
                            cex.i = 5, pch.i = 20, 
                             pch.j = 21,  text.cex.j =2, 
                             alpha.axes = 0.2,alpha.points.i = 1, axis1 = 1, axis2 = 2, title = " Asymmetrical plot for components 1 and 2")

alabels4MCA1 <- createxyLabels(resCA = resMCA.asym)

map.IJ.asym1 <- asymMap1$baseMap + asymMap1$I_points +
  asymMap1$J_labels + asymMap1$J_points + alabels4MCA1
print(map.IJ.asym1)

```

## Biplot for asymmetrical plot for component 2

Asymmetric plot deals with the attraction by the variables 
```{r}
asymMap2  <- createFactorMapIJ(resMCA.asym$ExPosition.Data$fi,resMCA.asym$ExPosition.Data$fj,
                             col.points.i = resMCA.asym$Plotting.Data$fi.col,
                             col.labels.j = resMCA.asym$Plotting.Data$fj.col,
                            cex.i = 5, pch.i = 20, 
                             pch.j = 21,  text.cex.j =2, 
                             alpha.axes = 0.2,alpha.points.i = 1, axis1 = 2, axis2 = 3, title = " Asymmetrical plot for components 2 and 3")

alabels4MCA2 <- createxyLabels(resCA = resMCA.asym)

map.IJ.asym2 <- asymMap2$baseMap + asymMap2$I_points +
  asymMap2$J_labels + asymMap2$J_points + alabels4MCA2
print(map.IJ.asym2)

```

## Biplot for asymmetrical plot for component 3

Asymmetric plot deals with the attraction by the variables 
```{r}
asymMap3  <- createFactorMapIJ(resMCA.asym$ExPosition.Data$fi,resMCA.asym$ExPosition.Data$fj,
                             col.points.i = resMCA.asym$Plotting.Data$fi.col,
                             col.labels.j = resMCA.asym$Plotting.Data$fj.col,
                            cex.i = 5, pch.i = 20, 
                             pch.j = 21,  text.cex.j =2, 
                             alpha.axes = 0.2,alpha.points.i = 1, axis1 = 3, axis2 = 4, title = " Asymmetrical plot for components 3 and 4")

alabels4MCA3 <- createxyLabels(resCA = resMCA.asym)

map.IJ.asym3 <- asymMap3$baseMap + asymMap3$I_points +
  asymMap3$J_labels + asymMap3$J_points + alabels4MCA3
print(map.IJ.asym3)

```

## Bootstrap Interval 

```{r}
BootCube.Gr <- Boot4Mean(resMCA.asym$ExPosition.Data$fi, 
                                    design = my_data$Department,
                                    niter = 100,
                                    suppressProgressBar = TRUE)
#_____________________________________________________________________
# Bootstrap ratios ----
bootRatios.Gr <- boot.ratio.test(BootCube.Gr$BootCube)
#*********************************************************************
# Mean Map
#  create the map for the means
#  get the means by groups

dataMeans <- getMeans(resMCA.asym$ExPosition.Data$fi, my_data$Department)
# a vector of color for the means
col4data <- resMCA.asym$Plotting.Data$fi.col
col4Means <- unique(col4data)
# the map
MapGroup <- createFactorMap(dataMeans,
                                       # use the constraint from the main map
                                       constraints = constraints.mca,
                                       col.points = col4Means,
                                       cex = 7,  # size of the dot (bigger)
                                       col.labels = col4Means,
                                       text.cex = 6)
# The map with observations and group means
a003.Map.I.withMeans <- a1 +
  MapGroup$zeMap_dots + MapGroup$zeMap_text
print(a003.Map.I.withMeans)
#_____________________________________________________________________
# Create the ellipses
# Bootstrapped CI ----
#_____________________________________________________________________
# Create Confidence Interval Plots
# use function MakeCIEllipses from package PTCA4CATA
GraphElli <- MakeCIEllipses(BootCube.Gr$BootCube[,1:2,],
                                       names.of.factors = c("Dimension 1","Dimension 2"),
                                       col = col4Means,
                                       p.level = .95
)
#_____________________________________________________________________
# create the I-map with Observations, means and confidence intervals
#
a004.Map.I.withCI <-  a1 + MapGroup$zeMap_text +  GraphElli
#_____________________________________________________________________
# plot it!
dev.new()
print(a004.Map.I.withCI)

```


## Contribution for variables 

```{r}
signed.ctrJ <- resMCA.sym$ExPosition.Data$cj * sign(resMCA.sym$ExPosition.Data$fj)
b003.ctrJ.s.1 <- PrettyBarPlot2(signed.ctrJ[,1],
                                threshold = 1 / NROW(signed.ctrJ),
                                font.size = 5,
                               # color4bar = gplots::col2hex(col4J.ibm), # we need hex code
                                main = 'MCA on the IBM-No-Attririon data Set: Variable Contributions (Signed)',
                                ylab = 'Contributions',
                                ylim = c(1.2*min(signed.ctrJ), 1.2*max(signed.ctrJ))
)
print(b003.ctrJ.s.1)
b004.ctrJ.s.2 <- PrettyBarPlot2(signed.ctrJ[,2],
                                threshold = 1 / NROW(signed.ctrJ),
                                font.size = 5,
                               # color4bar = gplots::col2hex(col4J.ibm), # we need hex code
                                main = 'MCA on the IBM-No-Attririon dataSet: Variable Contributions (Signed)',
                                ylab = 'Contributions',
                                ylim = c(1.2*min(signed.ctrJ), 1.2*max(signed.ctrJ))
)
print(b004.ctrJ.s.2)
b004.ctrJ.s.3 <- PrettyBarPlot2(signed.ctrJ[,3],
                                threshold = 1 / NROW(signed.ctrJ),
                                font.size = 5,
                               # color4bar = gplots::col2hex(col4J.ibm), # we need hex code
                                main = 'MCA on the IBM-No-Attririon dataSet: Variable Contributions (Signed)',
                                ylab = 'Contributions',
                                ylim = c(1.2*min(signed.ctrJ), 1.2*max(signed.ctrJ))
)
print(b004.ctrJ.s.3)

```

## Contribution for Rows

```{r}
signed.ctri <- resMCA.sym$ExPosition.Data$ci * sign(resMCA.sym$ExPosition.Data$fi)
b003.ctri.s.1 <- PrettyBarPlot2(signed.ctri[,1],
                                threshold = 1 / NROW(signed.ctri),
                                font.size = 5,
                               # color4bar = gplots::col2hex(col4J.ibm), # we need hex code
                                main = 'MCA on the IBM-No-Attririon data Set: Rows Contributions (Signed)',
                                ylab = 'Contributions',
                                ylim = c(1.2*min(signed.ctri), 1.2*max(signed.ctri)), horizontal = FALSE
)
print(b003.ctri.s.1)
b004.ctri.s.2 <- PrettyBarPlot2(signed.ctri[,2],
                                threshold = 1 / NROW(signed.ctri),
                                font.size = 5,
                               # color4bar = gplots::col2hex(col4J.ibm), # we need hex code
                                main = 'MCA on the IBM-No-Attririon dataSet: Rows Contributions (Signed)',
                                ylab = 'Contributions',
                                ylim = c(1.2*min(signed.ctri), 1.2*max(signed.ctri)) , horizontal = FALSE
)
print(b004.ctri.s.2)
b004.ctri.s.3 <- PrettyBarPlot2(signed.ctri[,3],
                                threshold = 1 / NROW(signed.ctri),
                                font.size = 5,
                               # color4bar = gplots::col2hex(col4J.ibm), # we need hex code
                                main = 'MCA on the IBM-No-Attririon dataSet: Rows Contributions (Signed)',
                                ylab = 'Contributions',
                                ylim = c(1.2*min(signed.ctri), 1.2*max(signed.ctri)),horizontal = FALSE
)
print(b004.ctri.s.3)

```


## Bootstrap Ratios for Variables 

```{r}
BR <- resMCA.inf$Inference.Data$fj.boots$tests$boot.ratios
laDim = 1
ba001.BR1 <- PrettyBarPlot2(BR[,laDim],
                            threshold = 2,
                            font.size = 5,
                            #color4bar = gplots::col2hex(col4J.ibm),
                            main = paste0( 'MCA on the IBM-NoAttrition data Set: Bootstrap ratio ',laDim),
                            ylab = 'Bootstrap ratios'
                            #ylim = c(1.2*min(BR[,laDim]), 1.2*max(BR[,laDim]))
)
print(ba001.BR1)
#
laDim = 2
ba002.BR2 <- PrettyBarPlot2(BR[,laDim],
                            threshold = 2,
                            font.size = 5,
                            #color4bar = gplots::col2hex(col4J.ibm),
                            main = paste0(
                              'MCA on the IBM-NoAttrition data Set: Bootstrap ratio ',laDim),
                            ylab = 'Bootstrap ratios'
)
print(ba002.BR2)
laDim = 3
ba002.BR3 <- PrettyBarPlot2(BR[,laDim],
                            threshold = 2,
                            font.size = 5,
                            main = paste0(
                              'MCA on the IBM-NoAttrition data Set: Bootstrap ratio ',laDim),
                            ylab = 'Bootstrap ratios'
)
print(ba002.BR3)


```

## Summary 











Questions to ask : 
Comparison between PCA nad MCA 
Asymmetrical plot factor map for i and j and biplot 




