---
title: "PLS"
author: "Richa Singh"
date: "10/25/2018"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
rm(list = ls())
graphics.off()
```

## PLS





## Dataset : IBM-HR-Employee-NoAttrition 

The dataset consists of 1233 observations and 32 variables describing the HR-IBM Employees. The variables Sub,Age,Monthly Income,Daily Rate,Hourly Rate, MonthlyRate,DistanceFromeHome,PerformanceRating, Education, JobInvolvement, Joblevel,StockOptionLevel,PercentSalaryHike,TotalWorkingYears,TrainingTimesLastYear,WorkLifeBalance,WorkLifeBalance,YearsAtCompany,YearsInCurrentRole,YearsSinceLastPromotion,YearsWithCurrManager,EnvironmentSatisfaction,JobSatisfaction,RelationshipSatisfaction,Attrition,BusinessTravel,EducationField,Gender,JobRole,MaritalStatus,OverTime are Qualitative variable.

```{r}
data <- read.csv("IBM-HR-Emplyee-NoAttrition.csv")
data1 <- data[,11:15] 
data2 <- data[,22:29] 
data2 <- data2[ -c(4) ]
data$Gender <- as.numeric(as.factor(data$Gender))
data$Department <- as.numeric(as.factor(data$Department))
data$JobRole <- as.numeric(as.factor(data$JobRole))
data$EducationField <- as.numeric(as.factor(data$EducationField))
design1 <- data$Gender
design2 <- data$Department
design3 <- data$JobRole
design4 <- data$EducationField
```

## Correlation Plot

```{r}
corrplot::corrplot(cor(data1,data2))
```

## Scree Plot

A Scree Plot is a simple line segment plot that shows the fraction of total variance in the data as explained or represented by each PC.```(In the PCA literature, the plot is called a 'Scree' Plot because it often looks like a 'scree' slope, where rocks have fallen down and accumulated on the side of a mountain.)```The scree plot shows the eigenvalues, the amount of information on each component. The number of components (the dimensionality of the factor space) is min(nrow(DATA), ncol(DATA)) minus 1.

```{r}
library(TExPosition)
resPLSC <- tepPLS(data1,data2,DESIGN = design1,graphs = FALSE)
```

```{r}
resPerm4PLSC <- perm4PLSC(data1, # First Data matrix 
                          data2, # Second Data matrix
                          nIter = 1000 # How mny iterations
                          )
```

```{r}
PlotScree(ev = resPLSC$TExPosition.Data$eigs, 
          p.ev = resPerm4PLSC$pEigenvalues,
          title = 'IBM-No-Attririon data Set. Eigenvalues Inference',
          plotKaiser = TRUE
)
```
## Permutation test for eigen-values

```{r}
zeDim = 1
pH1 <- prettyHist(
  distribution = resPerm4PLSC$permEigenvalues[,zeDim], 
           observed = resPerm4PLSC$pEigenvalues[zeDim], 
           xlim = c(-2, 2), # needs to be set by hand
           breaks = 20,
           border = "brown", 
           main = paste0("Permutation Test for Eigenvalue ",zeDim),
           xlab = paste0("Eigenvalue ",zeDim), 
           ylab = "", 
           counts = FALSE, 
           cutoffs = c( 0.975))
b005.PermTest <- recordPlot()
```
## Factor Maps for latent Variables lx vs ly for Gender type
```{r}
library(data4PCCAR)
library(PTCA4CATA)
library(corrplot)
library(ggplot2)
library(ExPosition)
library(InPosition)
constraints1 <- minmaxHelper(mat1 = resPLSC$TExPosition.Data$lx, mat2  = resPLSC$TExPosition.Data$ly)

df <- as.data.frame(cbind(resPLSC$TExPosition.Data$lx[,1],resPLSC$TExPosition.Data$ly[,1]))
dataMeans <- getMeans(df, design1)
col4data <- factor(design1)
col4Means <- c("#F8766D", "#00BFC4")
MapGroup <- createFactorMap(dataMeans,
                            # use the constraint from the main map
                            constraints = constraints1,
                            col.points = col4Means,
                            cex = 7,  # size of the dot (bigger)
                            col.labels = col4Means,
                            text.cex = 6)

GraphTI.Hull <- PTCA4CATA::MakeToleranceIntervals(df,
                                                  design = design1,
                                                  # line below is needed
                                                  names.of.factors =  c("Dim1","Dim2"), # needed 
                                                  col = col4Means,
                                                  line.size = .50, 
                                                  line.type = 3,
                                                  alpha.ellipse = .2,
                                                  alpha.line    = .4,
                                                  p.level       = .75)

z <- ggplot(data, aes(x=resPLSC$TExPosition.Data$lx[,1], y=resPLSC$TExPosition.Data$ly[,1])) + geom_point(aes(colour=col4data)) + MapGroup$zeMap_dots + MapGroup$zeMap_text + GraphTI.Hull
print(z)
```

```{r}
df1 <- as.data.frame(cbind(resPLSC$TExPosition.Data$lx[,2],resPLSC$TExPosition.Data$ly[,2]))
dataMeans1 <- getMeans(df1, design1)
col4data1 <- factor(design1)
col4Means1 <-  c("#F8766D", "#00BFC4")
MapGroup1 <- createFactorMap(dataMeans1,
                            # use the constraint from the main map
                            constraints = constraints1,
                            col.points = col4Means1,
                            cex = 7,  # size of the dot (bigger)
                            col.labels = col4Means1,
                            text.cex = 6)

GraphTI.Hull1 <- PTCA4CATA::MakeToleranceIntervals(df1,
                                                  design = design1,
                                                  # line below is needed
                                                  names.of.factors =  c("Dim1","Dim2"), # needed 
                                                  col = col4Means1,
                                                  line.size = .50, 
                                                  line.type = 3,
                                                  alpha.ellipse = .2,
                                                  alpha.line    = .4,
                                                  p.level       = .75)

z1 <- ggplot(data, aes(x=resPLSC$TExPosition.Data$lx[,2], y=resPLSC$TExPosition.Data$ly[,2])) + geom_point(aes(colour= col4data1)) + MapGroup1$zeMap_dots + MapGroup1$zeMap_text + GraphTI.Hull1
print(z1)
```

## Factor Maps for J

```{r}
baseMap.j2 <- createFactorMap(resPLSC$TExPosition.Data$fi,constraints = constraints,
                            col.points = resPLSC$Plotting.Data$fi.col, axis1 = 1, axis2 = 2,
                            cex = 1, pch = 20,
                            display.labels = TRUE
)
a2 <- baseMap.j2$zeMap + baseMap.j2$zeMap_dots 
print(a2)

```


```{r}
baseMap.j3 <- createFactorMap(resPLSC$TExPosition.Data$fj,constraints = constraints,
                            col.points = resPLSC$Plotting.Data$fj.col, axis1 = 1, axis2 = 2,
                            cex = 1, pch = 20,
                            display.labels = TRUE
)
a1 <- baseMap.j3$zeMap + baseMap.j3$zeMap_dots 
print(a1)

```
## Factor Map for latent variables lx vs ly for Departmenttype(1st Component)

```{r}
resPLSC1 <- tepPLS(data1,data2,DESIGN = design2,graphs = FALSE)
```

```{r}
constraints2 <- minmaxHelper(mat1 = resPLSC1$TExPosition.Data$lx, mat2  = resPLSC1$TExPosition.Data$ly)
df2 <- as.data.frame(cbind(resPLSC1$TExPosition.Data$lx[,1],resPLSC1$TExPosition.Data$ly[,1]))
dataMeans2 <- getMeans(df2, design2)
col4data2 <- factor(design2)
col4Means2 <-  c("#F8766D" ,"#00BA38" ,"#619CFF")
MapGroup2 <- createFactorMap(dataMeans2,
                            # use the constraint from the main map
                            constraints = constraints2,
                            col.points = col4Means2,
                            cex = 7,  # size of the dot (bigger)
                            col.labels = col4Means2,
                            text.cex = 6)
GraphTI.Hull2 <- PTCA4CATA::MakeToleranceIntervals(df2,
                                                  design = design2,
                                                  # line below is needed
                                                  names.of.factors =  c("Dim1","Dim2"), # needed 
                                                  col = col4Means2,
                                                  line.size = .50, 
                                                  line.type = 3,
                                                  alpha.ellipse = .2,
                                                  alpha.line    = .4,
                                                  p.level       = .75)



z2 <- ggplot(data, aes(x=resPLSC1$TExPosition.Data$lx[,1], y=resPLSC1$TExPosition.Data$ly[,1])) + geom_point(aes(colour = factor(design2))) + MapGroup2$zeMap_dots + MapGroup2$zeMap_text + GraphTI.Hull2 + xlab("lx 1st dimension") +
  ylab("ly 1st dimension") +
  ggtitle("lx vs ly for Department Type 1st dimension")
print(z2)
```

## Factor Map for latent variables lx vs ly for Departmenttype(2st Component)

```{r}
df3 <- as.data.frame(cbind(resPLSC1$TExPosition.Data$lx[,2],resPLSC1$TExPosition.Data$ly[,2]))
dataMeans3 <- getMeans(df3, design2)
col4data3 <- factor(design2)
col4Means3 <- c("#F8766D" ,"#00BA38" ,"#619CFF")
MapGroup3 <- createFactorMap(dataMeans3,
                            # use the constraint from the main map
                            constraints = constraints2,
                            col.points = col4Means3,
                            cex = 7,  # size of the dot (bigger)
                            col.labels = col4Means3,
                            text.cex = 6)
GraphTI.Hull3 <- PTCA4CATA::MakeToleranceIntervals(df3,
                                                  design = design2,
                                                  # line below is needed
                                                  names.of.factors =  c("Dim1","Dim2"), # needed 
                                                  col = col4Means3,
                                                  line.size = .50, 
                                                  line.type = 3,
                                                  alpha.ellipse = .2,
                                                  alpha.line    = .4,
                                                  p.level       = .75)


z3 <- ggplot(data, aes(x=resPLSC1$TExPosition.Data$lx[,2], y=resPLSC1$TExPosition.Data$ly[,2])) + geom_point(aes(colour = factor(design2))) + MapGroup3$zeMap_dots + MapGroup3$zeMap_text + GraphTI.Hull3 + xlab("lx 2nd dimension") +
  ylab("ly 2nd dimension") +
  ggtitle("lx vs ly for Department Type 2nd dimension")
print(z3)
```

## Factor Map for latent variables lx vs ly for JObRole(1st Component)

```{r}
resPLSC2 <- tepPLS(data1,data2,DESIGN = design3,graphs = FALSE)
```

```{r}
constraints3 <- minmaxHelper(mat1 = resPLSC2$TExPosition.Data$lx, mat2  = resPLSC2$TExPosition.Data$ly)
df4 <- as.data.frame(cbind(resPLSC2$TExPosition.Data$lx[,1],resPLSC2$TExPosition.Data$ly[,1]))
dataMeans4 <- getMeans(df4, design3)
col4data4 <- factor(design3)
col4Means4 <- c("#F8766D", "#D39200" ,"#93AA00" ,"#00BA38", "#00C19F", "#00B9E3", "#619CFF" ,"#DB72FB", "#FF61C3")
MapGroup4 <- createFactorMap(dataMeans4,
                            # use the constraint from the main map
                            constraints = constraints3,
                            col.points = col4Means4,
                            cex = 7,  # size of the dot (bigger)
                            col.labels = col4Means4,
                            text.cex = 6)
BootCube.Gr4 <- Boot4Mean(df4, 
                         design = design3,
                         niter = 100,
                         suppressProgressBar = TRUE)

#GraphElli4 <- MakeCIEllipses(BootCube.Gr4$BootCube,
#                            names.of.factors = c("Dimension 1","Dimension 2"),
#                            col = col4Means4 ,p.level = .95
#)

GraphTI.Hull4 <- PTCA4CATA::MakeToleranceIntervals(df4,
                                                  design = design3,
                                                  # line below is needed
                                                  names.of.factors =  c("Dim1","Dim2"), # needed 
                                                  col = col4Means4,
                                                  line.size = .50, 
                                                  line.type = 3,
                                                  alpha.ellipse = .2,
                                                  alpha.line    = .4,
                                                  p.level       = .75)

z4 <- ggplot(data, aes(x = resPLSC2$TExPosition.Data$lx[,1], y = resPLSC2$TExPosition.Data$ly[,1])) +
  geom_point(aes(color = col4data4)) + MapGroup4$zeMap_dots + MapGroup4$zeMap_text 
#+ GraphTI.Hull4 
#+ GraphElli4

print(z4)


```

## Factor Map for latent variables lx vs ly for JObRole(2nd Component)

```{r}
df5 <- as.data.frame(cbind(resPLSC2$TExPosition.Data$lx[,2],resPLSC2$TExPosition.Data$ly[,2]))
dataMeans5 <- getMeans(df5, design3)
col4data5 <- factor(design3)
col4Means5 <- c("#F8766D", "#D39200" ,"#93AA00" ,"#00BA38", "#00C19F", "#00B9E3", "#619CFF" ,"#DB72FB", "#FF61C3")
MapGroup5 <- createFactorMap(dataMeans5,
                            # use the constraint from the main map
                            constraints = constraints3,
                            col.points = col4Means5,
                            cex = 7,  # size of the dot (bigger)
                            col.labels = col4Means5,
                            text.cex = 6)
GraphTI.Hull5 <- PTCA4CATA::MakeToleranceIntervals(df5,
                                                  design = design3,
                                                  # line below is needed
                                                  names.of.factors =  c("Dim1","Dim2"), # needed 
                                                  col = col4Means5,
                                                  line.size = .50, 
                                                  line.type = 3,
                                                  alpha.ellipse = .2,
                                                  alpha.line    = .4,
                                                  p.level       = .75)


z5 <- ggplot(data, aes(x = resPLSC2$TExPosition.Data$lx[,2], y = resPLSC2$TExPosition.Data$ly[,2])) +
  geom_point(aes(color = factor(design3))) + MapGroup5$zeMap_dots + MapGroup5$zeMap_text + GraphTI.Hull5

print(z5)
```

## Factor Map for latent variables lx vs ly  for EducationField(1st Component)

```{r}
resPLSC3 <- tepPLS(data1,data2,DESIGN = design4,graphs = FALSE)
```

```{r}
constraints4 <- minmaxHelper(mat1 = resPLSC3$TExPosition.Data$lx, mat2  = resPLSC3$TExPosition.Data$ly)
df6 <- as.data.frame(cbind(resPLSC3$TExPosition.Data$lx[,1],resPLSC3$TExPosition.Data$ly[,1]))
dataMeans6 <- getMeans(df6, design4)
col4data6 <- factor(design4)
col4Means6 <- c("#F8766D" ,"#B79F00" ,"#00BA38" ,"#00BFC4" ,"#619CFF", "#F564E3")
MapGroup6 <- createFactorMap(dataMeans6,
                            # use the constraint from the main map
                            constraints = constraints4,
                            col.points = col4Means6,
                            cex = 7,  # size of the dot (bigger)
                            col.labels = col4Means6,
                            text.cex = 6)
GraphTI.Hull6 <- PTCA4CATA::MakeToleranceIntervals(df6,
                                                  design = design4,
                                                  # line below is needed
                                                  names.of.factors =  c("Dim1","Dim2"), # needed 
                                                  col = col4Means6,
                                                  line.size = .50, 
                                                  line.type = 3,
                                                  alpha.ellipse = .2,
                                                  alpha.line    = .4,
                                                  p.level       = .75)



z6 <- ggplot(data, aes(x=resPLSC3$TExPosition.Data$lx[,1], y=resPLSC3$TExPosition.Data$ly[,1])) + geom_point(aes(color = factor(design4))) + MapGroup6$zeMap_dots + MapGroup6$zeMap_text + GraphTI.Hull6
print(z6)
```

## Factor Map for latent variables lx vs ly  for EducationField(2nd Component)

```{r}
df7 <- as.data.frame(cbind(resPLSC3$TExPosition.Data$lx[,2],resPLSC3$TExPosition.Data$ly[,2]))
dataMeans7 <- getMeans(df7, design4)
col4data7 <- factor(design4)
col4Means7 <- c("#F8766D" ,"#B79F00" ,"#00BA38" ,"#00BFC4" ,"#619CFF", "#F564E3")
MapGroup7 <- createFactorMap(dataMeans7,
                            # use the constraint from the main map
                            constraints = constraints4,
                            col.points = col4Means7,
                            cex = 7,  # size of the dot (bigger)
                            col.labels = col4Means7,
                            text.cex = 6)

GraphTI.Hull7 <- PTCA4CATA::MakeToleranceIntervals(df6,
                                                  design = design4,
                                                  # line below is needed
                                                  names.of.factors =  c("Dim1","Dim2"), # needed 
                                                  col = col4Means6,
                                                  line.size = .50, 
                                                  line.type = 3,
                                                  alpha.ellipse = .2,
                                                  alpha.line    = .4,
                                                  p.level       = .75)


z7 <- ggplot(data, aes(x=resPLSC3$TExPosition.Data$lx[,2], y=resPLSC3$TExPosition.Data$ly[,2])) + geom_point(aes(color = factor(design4))) + MapGroup7$zeMap_dots + MapGroup7$zeMap_text+ GraphTI.Hull7

print(z7)
```


## Contributions for Variables 

```{r}
signed.ctrJ <- resPLSC$TExPosition.Data$cj * sign(resPLSC$TExPosition.Data$fj)
b003.ctrJ.s.1 <- PrettyBarPlot2(signed.ctrJ[,1],
                                threshold = 1 / NROW(signed.ctrJ),
                                font.size = 5,
                               # color4bar = gplots::col2hex(col4J.ibm), # we need hex code
                                main = 'IBM-No-Attririon data Set: Variable Contributions (Signed)',
                                ylab = 'Contributions',
                                ylim = c(1.2*min(signed.ctrJ), 1.2*max(signed.ctrJ)), horizontal = FALSE
)
print(b003.ctrJ.s.1)
b004.ctrJ.s.2 <- PrettyBarPlot2(signed.ctrJ[,2],
                                threshold = 1 / NROW(signed.ctrJ),
                                font.size = 5,
                               # color4bar = gplots::col2hex(col4J.ibm), # we need hex code
                                main = 'IBM-No-Attririon dataSet: Variable Contributions (Signed)',
                                ylab = 'Contributions',
                                ylim = c(1.2*min(signed.ctrJ), 1.2*max(signed.ctrJ)), horizontal = FALSE
)
print(b004.ctrJ.s.2)
b004.ctrJ.s.3 <- PrettyBarPlot2(signed.ctrJ[,3],
                                threshold = 1 / NROW(signed.ctrJ),
                                font.size = 5,
                               # color4bar = gplots::col2hex(col4J.ibm), # we need hex code
                                main = 'IBM-No-Attririon dataSet: Variable Contributions (Signed)',
                                ylab = 'Contributions',
                                ylim = c(1.2*min(signed.ctrJ), 1.2*max(signed.ctrJ)),horizontal = FALSE
)
print(b004.ctrJ.s.3)

```

## Contribution for Rows 

```{r}
signed.ctri <- resPLSC$TExPosition.Data$ci * sign(resPLSC$TExPosition.Data$fi)
b003.ctri.s.1 <- PrettyBarPlot2(signed.ctri[,1],
                                threshold = 1 / NROW(signed.ctri),
                                font.size = 5,
                               # color4bar = gplots::col2hex(col4J.ibm), # we need hex code
                                main = 'IBM-No-Attririon data Set: Rows Contributions (Signed)',
                                ylab = 'Contributions',
                                ylim = c(1.2*min(signed.ctri), 1.2*max(signed.ctri)), horizontal = FALSE
)
print(b003.ctri.s.1)
b004.ctri.s.2 <- PrettyBarPlot2(signed.ctri[,2],
                                threshold = 1 / NROW(signed.ctri),
                                font.size = 5,
                               # color4bar = gplots::col2hex(col4J.ibm), # we need hex code
                                main = 'IBM-No-Attririon dataSet: Rows Contributions (Signed)',
                                ylab = 'Contributions',
                                ylim = c(1.2*min(signed.ctri), 1.2*max(signed.ctri)) , horizontal = FALSE
)
print(b004.ctri.s.2)
b004.ctri.s.3 <- PrettyBarPlot2(signed.ctri[,3],
                                threshold = 1 / NROW(signed.ctri),
                                font.size = 5,
                               # color4bar = gplots::col2hex(col4J.ibm), # we need hex code
                                main = 'IBM-No-Attririon dataSet: Rows Contributions (Signed)',
                                ylab = 'Contributions',
                                ylim = c(1.2*min(signed.ctri), 1.2*max(signed.ctri)),horizontal = FALSE
)
print(b004.ctri.s.3)

```

## Load the file

```{r}
file4PLSCfunctions <- 'inferences4PLSC(4).R'
source(file4PLSCfunctions)
```

## Bootstrap Ratios for Variables 

```{r}
resBoot4PLSC <- Boot4PLSC(data1, # First Data matrix 
                          data2, # Second Data matrix
                          nIter = 1000, # How many iterations
                          Fi = resPLSC$TExPosition.Data$fi,
                          Fj = resPLSC$TExPosition.Data$fj,
                          nf2keep = 3,
                          critical.value = 2,
                          # To be implemented later
                          # has no effect currently
                          alphaLevel = .05)
print(resBoot4PLSC)
laDim = 1
ba001.BR1 <- PrettyBarPlot2(resBoot4PLSC$bootRatios.j[,laDim],
                            threshold = 2,
                            font.size = 5,
                            #color4bar = gplots::col2hex(col4J.ibm),
                            main = paste0( 'IBM-NoAttrition data Set: Bootstrap ratio ',laDim),
                            ylab = 'Bootstrap ratios'
                            #ylim = c(1.2*min(BR[,laDim]), 1.2*max(BR[,laDim]))
)
print(ba001.BR1)
#
laDim = 2
ba002.BR2 <- PrettyBarPlot2(resBoot4PLSC$bootRatios.j[,laDim],
                            threshold = 2,
                            font.size = 5,
                            #color4bar = gplots::col2hex(col4J.ibm),
                            main = paste0(
                              'IBM-NoAttrition data Set: Bootstrap ratio ',laDim),
                            ylab = 'Bootstrap ratios'
)
print(ba002.BR2)
laDim = 3
ba002.BR3 <- PrettyBarPlot2(resBoot4PLSC$bootRatios.j[,laDim],
                            threshold = 2,
                            font.size = 5,
                            main = paste0(
                              'IBM-NoAttrition data Set: Bootstrap ratio ',laDim),
                            ylab = 'Bootstrap ratios'
)
print(ba002.BR3)
```
## Bootstrap Ratios for Rows

```{r}
laDim = 1
ba001.BR11 <- PrettyBarPlot2(resBoot4PLSC$bootRatios.i[,laDim],
                            threshold = 2,
                            font.size = 5,
                            #color4bar = gplots::col2hex(col4J.ibm),
                            main = paste0( 'IBM-NoAttrition data Set: Bootstrap ratio ',laDim),
                            ylab = 'Bootstrap ratios'
                            #ylim = c(1.2*min(BR[,laDim]), 1.2*max(BR[,laDim]))
)
print(ba001.BR11)
#
laDim = 2
ba002.BR21 <- PrettyBarPlot2(resBoot4PLSC$bootRatios.i[,laDim],
                            threshold = 2,
                            font.size = 5,
                            #color4bar = gplots::col2hex(col4J.ibm),
                            main = paste0(
                              'IBM-NoAttrition data Set: Bootstrap ratio ',laDim),
                            ylab = 'Bootstrap ratios'
)
print(ba002.BR21)
laDim = 3
ba002.BR31 <- PrettyBarPlot2(resBoot4PLSC$bootRatios.i[,laDim],
                            threshold = 2,
                            font.size = 5,
                            main = paste0(
                              'IBM-NoAttrition data Set: Bootstrap ratio ',laDim),
                            ylab = 'Bootstrap ratios'
)
print(ba002.BR31)
```


## Summary 






