---
title: "BADA"
author: "Richa Singh"
date: "11/9/2018"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
rm(list = ls())
graphics.off()
```

## BADA

Barycentric discriminant analysis (BADA) is a robust version of discriminant analysis that is used to assign, to pre-defined groups (also called categories), observations described by multiple variables.The goal of BADA is to combine the measurements to create new variables (called components or discriminant variables) that best separate the categories. These discriminant variables are also used to assign the original observations or ???new??? observations to the a-priori defined categories.Barycentric discriminant analysis is a robust version of discriminant analysis that is used when multiple measurements describe a set of observations in which each observation be- longs to one category (i.e., group) from a set of a priori defined categories. BADA combines the original variables to create new variables that best separate the groups and that can also be used to optimally assign old or new observations to these categories. The quality of the performance is evaluated by cross-validation techniques that estimate the performance of the classification model for new observations.
BADA is a very versatile technique that can be declined in several different varieties that can handle, for example, qualitative data and data structured in blocks. This versatility make BADA particularly suited for the analysis of multi-modal and Big data.

## Dataset 

The dataset consists of 1233 observations and 32 variables describing the HR-IBM Employees. 
The variables:
Quantitative Variables :
Sub,Age,Monthly Income,Daily Rate,Hourly Rate, MonthlyRate, DistanceFromeHome,PercentSalaryHike, TotalWorkingYears,TrainingTimesLastYear,WorkLifeBalance,WorkLifeBalance,YearsAtCompany,YearsInCurrentRole,YearsSinceLastPromotion,YearsWithCurrManager
Ordinal Variables : 
PerformanceRating, Education, JobInvolvement,Joblevel,StockOptionLevel,EnvironmentSatisfaction, JobSatisfaction,RelationshipSatisfaction
Qualitative variable :
Attrition,BusinessTravel,EducationField,Gender,JobRole,MaritalStatus,OverTime
```{r include=FALSE}
data <- read.csv("IBM-HR-Emplyee-NoAttrition.csv")
rownames(data) <- data$Subj
data1 <- data[sapply(data, function(x) !is.factor(x))]
data1 <- subset(data1, select=-Subj)
```

```{r include=FALSE}
(library(Matrix))
(library(prettyGraphs)) 
(library(ExPosition))
(library(InPosition)) 
(library(TExPosition))
(library(TInPosition))
(library(PTCA4CATA)) 
(library(data4PCCAR))
(library(dplyr))
(library(gridExtra))
(library(grid))
```


```{r}
datae <- data[,8]
resBADA <- tepBADA(DATA = data1,
                   scale = 'SS1', center = TRUE,
                   DESIGN = datae,
                   make_design_nominal = TRUE,
                   group.masses = NULL,
                   weights = NULL, graphs =  FALSE)
```

## ScreePlot

```{r}
PlotScree(ev = resBADA$TExPosition.Data$eigs,
          p.ev = NULL, max.ev = NULL, alpha = 0.05,
          col.ns = "#006D2C", col.sig = "#54278F",
          title = "Explained Variance per Dimension",plotKaiser = TRUE)
```

## HeatMap

```{r}
color4Var <- prettyGraphs::prettyGraphsColorSelection(ncol(data1))
data2 <- makeNominalData(as.matrix(data[,8]))
corrplot::corrplot(cor(data2,data1))
```

```{r}
Fk <- resBADA$TExPosition.Data$fi
Fi <- resBADA$TExPosition.Data$fii
Fj <- resBADA$TExPosition.Data$fj
```

## Factor Map J

```{r}
col4Var <- prettyGraphsColorSelection(NCOL(data1))
baseMap.j <- PTCA4CATA::createFactorMap(Fj,
                            col.points   = col4Var,
                            alpha.points =  .3,
                            col.labels   = col4Var)
# A graph for the J-set
aggMap.j <- baseMap.j$zeMap_background + # background layer
  baseMap.j$zeMap_dots + baseMap.j$zeMap_text # dots & labels
# We print this Map with the following code
dev.new()
print(aggMap.j)
zeLines <- ggplot2::annotate("segment", x = c(0), y = c(0),
                    xend = Fj[,1],
                    yend = Fj[,2],
                    color = col4Var,
                    alpha = .5,
                    arrow = arrow(length = unit(.3, "cm") ) )
# Create the map by adding background, labels, and arrows:
aggMap.j.arrows <- baseMap.j$zeMap_background +
                                      zeLines + baseMap.j$zeMap_text
dev.new()
print(aggMap.j.arrows)
```

## Factor Map I

```{r}
baseMap.i <- PTCA4CATA::createFactorMap(Fi,
                                        col.points   = resBADA$Plotting.Data$fii.col,
                                        alpha.points =  .3)
# Plain map with color for the I-set
aggMap.i <- baseMap.i$zeMap_background + baseMap.i$zeMap_dots
#---------------------------------------------------------------------
# print this Map
dev.new()
print(aggMap.i)
```

```{r}
col4data <- resBADA$Plotting.Data$fii.col
col4Means <- unique(col4data)
# create the map for the means
MapGroup    <- PTCA4CATA::createFactorMap(Fk,
                                axis1 = 1, axis2 = 2,
                                constraints = baseMap.i$constraints,
                                title = NULL,
                                col.points = col4Means,
                                display.points = TRUE,
                                pch = 19, cex = 5,
                                display.labels = TRUE,
                                col.labels = col4Means,
                                text.cex = 4,
                                font.face = "bold",
                                font.family = "sans",
                                col.axes = "darkorchid",
                                alpha.axes = 0.2,
                                width.axes = 1.1,
                            col.background = adjustcolor("lavender",
                                                      alpha.f = 0.2),
                            force = 1, segment.size = 0)
# The map with observations and group means
aggMap.i.withMeans <- aggMap.i+
  MapGroup$zeMap_dots + MapGroup$zeMap_text
#---------------------------------------------------------------------
# plot it!
dev.new()
print(aggMap.i.withMeans)
```

## Create 75% Tolerance interval polygons
```{r}
GraphTI.Hull.90 <- MakeToleranceIntervals(Fi,
                                  as.factor(datae),
                                  names.of.factors = c("Dim1","Dim2"),
                                  col = unique(col4data),
                                  line.size = .5, line.type = 3,
                                  alpha.ellipse = .2,
                                  alpha.line = .4,
                                  p.level = .75, # 75% TI
                                  type = 'hull' #
                                          # use 'hull' for convex hull
)
#---------------------------------------------------------------------
# Create the map
aggMap.i.withHull <- aggMap.i +
  GraphTI.Hull.90 + MapGroup$zeMap_dots +
  MapGroup$zeMap_text +  MapGroup$zeMap_dots
#---------------------------------------------------------------------
# Plot it!
dev.new()
print(aggMap.i.withHull)
```


## Inferences
```{r}
resBADA.inf <- tepBADA.inference.battery(DATA = data1,
                   scale = 'SS1', center = TRUE,
                   DESIGN = datae,
                   make_design_nominal = TRUE,
                   group.masses = NULL,
                   weights = NULL,
                   graphs = FALSE,
                   k = 2,
                   test.iters = 100,
                   critical.value = 2)
#---------------------------------------------------------------------
# Confusion matrices
# To be saved as table
fixedCM   <-   resBADA.inf$Inference.Data$loo.data$fixed.confuse
looedCM   <- resBADA.inf$Inference.Data$loo.data$loo.confuse

#---------------------------------------------------------------------
# Create Confidence Interval Plots
BootCube <- resBADA.inf$Inference.Data$boot.data$fi.boot.data$boots
dimnames(BootCube)[[2]] <- c("Dimension 1","Dimension 2")
# use function MakeCIEllipses from package PTCA4CATA
GraphElli <- MakeCIEllipses(BootCube[,1:2,],
                  names.of.factors = c("Dimension 1","Dimension 2"),
                  col = col4Means,
                  p.level = .95
)
#---------------------------------------------------------------------
# create the I-map with Observations, means and confidence intervals
#
aggMap.i.withCI <-  aggMap.i +  GraphElli + MapGroup$zeMap_text
#---------------------------------------------------------------------
# plot it!
dev.new()
print(aggMap.i.withCI)
```

## Contribution 

```{r}
signed.ctrJ1 <- resBADA$TExPosition.Data$cj * sign(resBADA$TExPosition.Data$fj)
b003.ctrJ.s.11 <- PrettyBarPlot2(signed.ctrJ1[,1],
                                threshold = 1 / NROW(signed.ctrJ1),
                                font.size = 5,
                                # color4bar = gplots::col2hex(col4J.ibm), # we need hex code
                                main = 'BADA on the IBM-No-Attririon data Set: Variable Contributions (Signed)',
                                ylab = 'Contributions',
                                ylim = c(1.2*min(signed.ctrJ1), 1.2*max(signed.ctrJ1))
)
print(b003.ctrJ.s.11)
b004.ctrJ.s.21 <- PrettyBarPlot2(signed.ctrJ1[,2],
                                threshold = 1 / NROW(signed.ctrJ1),
                                font.size = 5,
                                # color4bar = gplots::col2hex(col4J.ibm), # we need hex code
                                main = 'BADA on the IBM-No-Attririon dataSet: Variable Contributions (Signed)',
                                ylab = 'Contributions',
                                ylim = c(1.2*min(signed.ctrJ1), 1.2*max(signed.ctrJ1))
)
print(b004.ctrJ.s.21)
```

## Correlation Circle

```{r}
correlationCircle <- correlationPlotter(data_matrix = data1 , factor_scores = resBADA$TExPosition.Data$lx , x_axis = 1, y_axis = 2, col = NULL ,pch = NULL, xlab = NULL, ylab = NULL, main = "Correlation Circle", asp = 1, dev.new = FALSE) 
```

## Bootstrap Ratios 

```{r}
BR1 <- resBADA.inf$Inference.Data$boot.data$fj.boot.data$tests$boot.ratios
laDim = 1
ba001.BR11 <- PrettyBarPlot2(BR1[,laDim],
                            threshold = 2,
                            font.size = 5,
                            #color4bar = gplots::col2hex(col4J.ibm),
                            main = paste0( 'BADA on the IBM-NoAttrition data Set: Bootstrap ratio ',laDim),
                            ylab = 'Bootstrap ratios'
                            #ylim = c(1.2*min(BR[,laDim]), 1.2*max(BR[,laDim]))
)
print(ba001.BR11)
#
laDim = 2
ba002.BR21 <- PrettyBarPlot2(BR1[,laDim],
                            threshold = 2,
                            font.size = 5,
                            #color4bar = gplots::col2hex(col4J.ibm),
                            main = paste0(
                              'BADA on the IBM-NoAttrition data Set: Bootstrap ratio ',laDim),
                            ylab = 'Bootstrap ratios'
)
print(ba002.BR21)
```

## Random (LOO) confusion matrix

```{r}
cm <- as.matrix(resBADA$TExPosition.Data$assign$confusion)
correctly_classified <- sum(diag(cm))
Incorrectly_classified <- sum(cm)-sum(diag(cm))
print(sum(diag(cm))/sum(cm))
```

## Summary 

