---
title: "CookbookTemplate_v3"
author: "Michael A. Kriegsman"
date: "September 18, 2017"
output:
  word_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ExPosition)
```

# Method: PCA
Principal component analysis (PCA) is... 


## Data set: state.x77
State.x77 is... 

```{r data_set}
head(state.x77)
```

Create DESIGN variables.

Below, I show you how to work with DESIGN variables and colors. Please Delete our comments, and write your own comments in here.

```{r DESIGN}
#Create an empty list, called DESIGN
DESIGN <- list()

#Let's set the group names (labels), and establish which row belongs to which group (as a DESIGN vector, and then as a DESIGN matrix)
#For now, let's work on a DESIGN on the rows. The DESIGN vairable is called region. 
#The group names are labeled:
DESIGN$rows$Region$labels <- unique(state.region) #to see what's here, highlight and run "state.region". Then also highlight and run "unique(state.region)". 

#This variable is pretty easy... the DESIGN vector (a vector that describes which row belongs to which group) is given by state.region
DESIGN$rows$Region$vec <- state.region

#Convert the vector to a matrix
DESIGN$rows$Region$mat <- makeNominalData(as.matrix(DESIGN$rows$Region$vec))
#notice that makeNominalData wants to have a matrix, so we use as.matrix(DESIGN_vector)
```


We also want to work with the colors of the groups.


```{r colors_1}
#One way to pick colors is automatically, like how epPCA does it...
automatic_colors <- createColorVectorsByDesign(DESIGN$rows$Region$mat)
#notice that automatic_colors$oc gives the colors for each observation, and automatic_colors$gc gives the colors for each group.
```

```{r colors_2}

###Make your own colors, and write your own comments ###

#But, instead of being automatic, we often want to select our colors by hand. 
#Let's just pick 4 colors (you should choose your own)
DESIGN$rows$Region$color_groups <- c("red", "green", "goldenrod", "#305ABF")

#and let's just look at the colors we've picked, to be sure they're pretty enough
plot(c(1:4), pch=15, cex=4, col=DESIGN$rows$Region$color_groups)
#Actually, these are quite ugly, but I'll let you pick your own pretty colors. 
#But don't show this plot in your cookbook, this is just for you to select colors. 

#Now, we need to create a vector of the group colors, to show which row belongs to which group.
#One group at a time, let's trade the group names for the color names... 
### Follow me slowly, and run each line to see what's inside

#First, copy the group names
DESIGN$rows$Region$color_observ <- as.matrix(DESIGN$rows$Region$vec) #we want to change the factor to a matrix, because they work differently

#Then, for each group, replace the group name with the group's color
DESIGN$rows$Region$color_observ[which(DESIGN$rows$Region$vec=="South")]         <- DESIGN$rows$Region$color_groups[1]
DESIGN$rows$Region$color_observ[which(DESIGN$rows$Region$vec=="West")]          <- DESIGN$rows$Region$color_groups[2]
DESIGN$rows$Region$color_observ[which(DESIGN$rows$Region$vec=="Northeast")]     <- DESIGN$rows$Region$color_groups[3]
DESIGN$rows$Region$color_observ[which(DESIGN$rows$Region$vec=="North Central")] <- DESIGN$rows$Region$color_groups[4]
#Be sure you get the group names right, and keep track of which group is which, so you can create your legend!
```

## Results
Center? Scale?

```{r analyze, echo = TRUE}
res_pca <- epPCA(state.x77, center = TRUE, scale = TRUE, graphs = FALSE)
```


## Plots

PCA gives... THESE plots...

### Scree Plot


```{r scree plot}
name_the_scree <- plot(res_pca$ExPosition.Data$eigs,
                       ylab = "Eigenvalues",
                       xlab = "Compnents",
                       type = "l",
                       main = "",
                       ylim = c(-1, max(res_pca$ExPosition.Data$eigs)))
points(res_pca$ExPosition.Data$eigs, cex = 2, pch = 19, col = "darkorchid4")
points(res_pca$ExPosition.Data$eigs, cex = 2, pch = 21, col = "black")

```

### Factor scores
Factor scores are the coordinates of the 50 states on the components. The distances between them show which states are most similar. Factor scores (states) can be color-coded to help interpret the components.

You may or may not want to leave display_names=TRUE... 

```{r factor scores}
name_the_plot <- prettyPlot(data_matrix = res_pca$ExPosition.Data$fi,  
                            dev.new=FALSE,
                            main = "State.x77 Row Factor Scores",
                            x_axis = 1, y_axis = 2, 
                            contributionCircles = FALSE, contributions = res_pca$ExPosition.Data$ci, 
                            display_points = TRUE, pch = 21, cex = 1.2, col = DESIGN$rows$Region$color_observ, 
                            display_names = TRUE, 
                            xlab = paste0("Component 1 Inertia: ", round(res_pca$ExPosition.Data$t[1],3), "%"),
                            ylab = paste0("Component 2 Inertia: ", round(res_pca$ExPosition.Data$t[2],3), "%")
                            )

legend(x="topright", pch = 15, legend = DESIGN$rows$Region$labels, col=DESIGN$rows$Region$color_groups)
```







* Component 1: 

* Component 2: 

* Component 3? Include whatever is relevant.

### Loadings
Loadings are...

```{r}
name_another_plot <- prettyPlot(data_matrix = res_pca$ExPosition.Data$fj,  
                            dev.new=FALSE,
                            main = "State.x77 Column Loadings",
                            x_axis = 1, y_axis = 2, 
                            contributionCircles = FALSE, contributions = res_pca$ExPosition.Data$cj, 
                            display_points = TRUE, pch = 21, cex = 1.2, col = res_pca$Plotting.Data$fj.col, 
                            display_names = TRUE, 
                            xlab = paste0("Component 1 Inertia: ", round(res_pca$ExPosition.Data$t[1],3), "%"),
                            ylab = paste0("Component 2 Inertia: ", round(res_pca$ExPosition.Data$t[2],3), "%")
                            )
```

* Component 1: 

* Component 2: 

* Component 3? Include whatever is relevant.

### Include other relevant plots...


##Conclusions
